---
title: "Patterns & Trends Project Work"
author: "Ramon und Dimitri"
format:  
  html:
    toc: true
    toc-location: left
    page-layout: full
    css: style.css
    number-sections: true
    number-tables: true
    cap-location: top

execute:
  echo: false
  fig-align: center
  fig-width: 12
  fig-height: 7
bibliography: bibliography.bib
---



# Abstract {.unnumbered .unlisted}

This study analyzes the movement patterns of two students using Google Timeline tracking and SimaPro for GHG emissions assessment. By segmenting rest and movement phases, identifying key locations, and comparing transportation modes, the project quantifies mobility behaviors and environmental impact. The findings will be visualized through maps and statistical comparisons, ensuring an insightful sustainability analysis. <span style="color:red;">Vergleich mit der Schweiz erwähnen</span>

<br><br>

# Introduction

In recent years, the tracking and analysis of human mobility have gained increasing importance—both for understanding movement behavior and for assessing its environmental impact. The growing availability of spatial and temporal data from digital devices presents new opportunities to link individual travel patterns with greenhouse gas (GHG) emissions, thereby contributing to broader climate change research.

Transport plays a key role in global GHG emissions and is one of the fastest-growing sources of climate-relevant emissions. In 2019, the transport sector emitted approximately 8.7 to 8.9 gigatonnes of CO₂-equivalent (GtCO₂-eq), a significant increase from 5.0 to 5.1 GtCO₂-eq in 1990. This growth positions the transport sector as the fourth-largest contributor to global GHG emissions, following the power, industry, and agriculture, forestry and land use (AFOLU) sectors. Transport accounts for around 15% of total global GHG emissions and approximately 23% of energy-related CO₂ emissions. Since 2010, emissions from the sector have grown at an average annual rate of 1.8%, making it the fastest-growing end-use sector [@IPCC_2022_WGIII_Ch_10].

The structure of global transport emissions reveals a sector heavily reliant on fossil fuels, especially for road-based mobility. As of 2019, road transport was by far the largest contributor, responsible for approximately 6.1 GtCO₂-eq, or about 69–70% of total transport emissions. This includes both passenger vehicles and freight transport, such as trucks and vans. The dominance of road transport underscores its central role in global mobility systems and its dependence on internal combustion engine vehicles [@IPCC_2022_WGIII_Ch_10].

International shipping is the second-largest source of transport emissions, contributing around 0.8 GtCO₂-eq (9–11% of the sector’s total). It is vital for global trade, with long-distance shipping routes forming the backbone of international logistics. Close behind is international aviation, which accounted for approximately 0.6 GtCO₂-eq, or 7–12% of transport emissions. The rise in air travel—particularly long-haul international flights—has made aviation one of the fastest-growing sources of emissions. Between 2010 and 2019, emissions from international aviation increased at an average annual rate of 3.4%, driven by expanding global mobility, tourism, and business travel [@IPCC_2022_WGIII_Ch_10].

In contrast, rail transport contributes only about 1% of direct emissions. Rail is typically more energy-efficient and lower in emissions, especially where electrified networks are in use. However, its limited share of global passenger and freight activity means its overall emissions impact remains relatively small. The remaining emissions—around 1.4 GtCO₂-eq or 16% of the sector total—come from various sources including domestic aviation, buses, and other forms of collective or niche transport. While each of these modes contributes less individually, they still offer significant potential for emissions reduction through cleaner fuels, increased efficiency, and modal shifts [@IPCC_2022_WGIII_Ch_10].

Moreover, detailed emission inventories help translate mobility behavior into CO₂ equivalents. In Switzerland, for example, approximately 13 tonnes of CO₂-equivalent are emitted per capita annually, including both domestic and abroad emissions [@bafu2025klima]. Of this total, only around 4 tonnes occur within Switzerland, while the remaining emissions are generated abroad, primarily due to the import of goods and services. About one-third of these emissions stem from mobility, excluding international flights [@bafu2025treibhausgas]. This distinction is important when assessing responsibility and reduction potential, as a significant portion of the environmental impact is embedded in global supply chains and consumption patterns [@bafu2025wirtschaftskonsum]. Given Switzerland’s population of approximately 8.8 million (as of 2023), this amounts to an estimated 114 million tonnes of CO₂-equivalent emissions per year.

@nathani2022umwelt estimate that mobility behavior in Switzerland leads to at least 2.4 tonnes of CO₂e per capita, with over 1.5 tonnes resulting from domestic direct emissions—such as burning fuel in vehicle engines. As such, transport is the sector responsible for the largest share of GHG emissions. The United States presents a similar picture, where transport also ranks as the top emitting sector [@us2019us].

Without significant intervention, transport-related emissions are projected to increase by 16–50% by 2050. However, in scenarios aligned with the 1.5°C climate target, a 59% reduction in transport CO₂e emissions by 2050 (compared to 2020 levels) is required. More ambitious pathways, such as the “1.5°C Low Demand” scenario, envision reductions of up to 90% [@IPCC_2022_WGIII_Ch_10].

This study is grounded in Computational Movement Analysis (CMA), which provides the theoretical and methodological framework. As @laube2014computational explains, CMA integrates geographic information science, computer science, and statistics to analyze movement through space and time. Geographic information science offers spatial modeling and spatio-temporal operations; computer science supports data handling, querying, and analysis via algorithms and simulations; and statistics contribute methods for describing and modeling dynamic movement patterns.

<br><br>

## Research objective

This research investigates the mobility patterns of two students, focusing on how their travel behavior differs in terms of movement, rest, visited locations, and transportation choices. Building on their Google Timeline data, we aim to quantify the environmental footprint of their mobility and compare it to national benchmarks. This analysis not only sheds light on personal movement dynamics but also demonstrates how CMA and lifecycle assessment tools like Simapro and Ecoinvent can be integrated to estimate the GHG impact of individual mobility.

In this work the google maps data from Ramon and Dimitri will be analyzed. The data is stored in a json file. The data contains the GPS coordinates of the two persons. The data will be visualized on a map.
Further, the data will be analyzed and the distance between the two persons will be calculated. <span style="color:red;">Umschreiben</span>

<br><br>

## Research Questions

1.	How do the movement patterns of the two students differ?

    a.	What is the ratio between rest and movement? (Segmentation) 
    
    b.	Which are the most frequently visited locations?
    
    c.	How does the use of transportation modes differ between the samples over a specific period?
    
    d.	What impact do the movement patterns have on each student's CO₂e footprint?
    
    e.  How do the extrapolated emissions of the students perform in comparison of the Swiss average?

<br><br>

# Methodology & Data 
    
This chapter outlines the methodological framework and data sources used to analyze individual mobility patterns and their associated greenhouse gas (GHG) emissions. The analysis combines high-resolution spatio-temporal data from Google Timeline with environmental impact modeling using life cycle assessment (LCA) tools. Computational Movement Analysis (CMA) principles serve as the theoretical foundation for structuring and interpreting mobility trajectories. Data preprocessing and spatial analysis were conducted in R, using a suite of packages for JSON parsing, spatial visualization, and statistical modeling. Together, these approaches enable a comprehensive assessment of movement behavior and its environmental implications at both local and international scales.

<br><br>

## R, Quarto and R packages

For this analysis, R [@r2024] was used for data cleaning, spatial analysis, visualization, and statistical modeling. The code was developed and executed using RStudio [@rstudio2024], an integrated development environment that facilitates reproducible research. The report itself was written using Quarto [@quarto2024], an open-source scientific and technical publishing system that supports dynamic documents and integrates seamlessly with R and RStudio.

```{r, warning=FALSE, message=FALSE, echo=TRUE}
library(jsonlite)     # For parsing and generating JSON data.
library(dplyr)        # For data manipulation using a grammar of data transformation.
library(tidyr)        # For tidying and reshaping data into a consistent format.
library(sf)           # For handling and analyzing spatial (vector) data using simple features.
library(ggplot2)      # For creating static graphics and visualizations using the grammar of graphics.
library(tmap)         # For creating thematic maps, both static and interactive.
library(tidyverse)    # A collection of R packages for data science, including ggplot2, dplyr, tidyr, etc.
library(leaflet)      # For creating interactive web maps using the Leaflet JavaScript library.
library(purrr)        # For functional programming, especially working with lists and iterating functions.
library(webshot2)     # For taking screenshots of web pages or HTML widgets.
library(htmlwidgets)  # For embedding JavaScript widgets (e.g., interactive plots/maps) in R Markdown or Shiny apps.
library(ggrepel)      # For improving the readability of ggplot2 text labels by avoiding overlap.
library(patchwork)    # For combining multiple ggplot2 plots into a single layout.
library(knitr)        # For dynamic report generation with R Markdown and LaTeX/HTML/Word output.
library(kableExtra)   # For enhancing tables created with knitr::kable, especially in HTML/PDF reports.
library(gt)           # For creating elegant, customizable tables for reporting.
library(RColorBrewer) # For generating color palettes for plots, especially categorical or sequential color scales.

```

<br><br>

## Google Timeline

Google Timeline data [@google2025timeline] was used as the primary source for capturing individual mobility patterns. This data is collected via Google Location History and is available to users who have enabled location tracking on their mobile devices. It offers high-resolution spatio-temporal data (see @tbl-google-timeline-data).

```{r, echo=FALSE}
#| label: tbl-google-timeline-data
#| tbl-cap: Overview of data elements available from Google Timeline exports [@google2025timeline].

timeline_data <- data.frame(
  `Data Element` = c(
    "Timestamp (start and end)",
    "Latitude and longitude",
    "Place name and address",
    "Activity type (e.g., visit, travel)",
    "Transportation mode",
    "Travel distance",
    "Dwell time",
    "Confidence level (Google's certainty rating)",
    "Device information (optional)",
    "Semantic place type (e.g., 'restaurant', 'university')"
  ),
  `Description` = c(
    "Time a location was entered and exited",
    "Precise GPS coordinates of locations and routes",
    "Named location or address from Google's POI database",
    "Distinguishes between staying at a place or being in transit",
    "Inferred travel mode such as walking, driving, cycling, bus, etc.",
    "Distance travelled per segment (estimated by Google)",
    "Duration spent at a specific location",
    "Google's internal confidence in its mode or place inference",
    "Device type or app used to record location, if available",
    "High-level category assigned to places visited"
  )
)

kable(timeline_data)
```

<br><br>

## Spatial Data analysis (CMA) 

The analysis of spatio-temporal mobility data in this study is grounded in core principles from computational movement analysis. The preprocessing and segmentation of individual movement trajectories was based on the framework described by @laube2014computational, particularly the methods outlined in Chapter 3.2.1: Segmentation and Filtering. Raw data from Google Location Services were organized into structured trajectories by leveraging the temporal and categorical information provided in the JSON files, including timestamps, transport mode labels, and location types.

The movement records were treated as event-based sequences, segmented into daily trip chains using changes in transportation mode and temporal gaps between recorded segments [@laube2014computational]. This segmentation process followed the conceptual structure proposed by @dodge2008taxonomy, who introduced a taxonomy of movement patterns that supports interpretation of temporal regularity and spatial extent—useful for distinguishing between local and long-distance activities.

<br><br>

## Life Cycle Data

<br><br>

### SimaPro

SimaPro 10.1 [@simapro2025] was used as the primary life cycle assessment (LCA) software for modeling and calculating greenhouse gas (GHG) emissions associated with transportation activities. It offers a robust platform for quantifying environmental impacts across full life cycles, including direct and indirect emissions related to fuel production, vehicle operation, and infrastructure. The software enables transparent modeling, flexible scenario comparisons, and integration of context-specific activity data, making it well-suited for mobility-related carbon footprint analyses.

<br><br>

### Ecoinvent

The Ecoinvent v3.11 database [@ecoinvent2024] was used as the life cycle inventory (LCI) source within SimaPro. Ecoinvent provides high-quality, peer-reviewed datasets for a wide range of economic activities, including detailed data on transport processes. For this study, specific datasets were selected to model emissions from various transportation modes—such as passenger cars (internal combustion engine and electric), buses, regional trains, trams, bicycles, and freight ferries. These datasets correspond to full process names from the Ecoinvent database, ensuring transparency and reproducibility in the modeling approach (see @tbl-transport-modes). The selection was based on the transport types and distances identified through Google Timeline tracking.

The LCI data include upstream and downstream processes, enabling a comprehensive cradle-to-grave emissions analysis. This approach allows for a detailed estimation of GHG emissions associated with each mode of transportation used during the analyzed period.

```{r, echo=FALSE}
#| label: tbl-transport-modes
#| tbl-cap: Ecoinvent transport processes used in the LCA model [@ecoinvent2024].


transport_data <- data.frame(
  `Transport Mode` = c(
    "Train",
    "Car (Internal Combustion Engine)",
    "Car (Electric)",
    "Motorcycle",
    "Bicycle",
    "Bus",
    "Tram",
    "Ferry"
  ),
  `Ecoinvent Process Name` = c(
    "Transport, passenger train {CH}| transport, passenger train, regional | Cut-off, U",
    "Transport, passenger car with internal combustion engine {RER}| transport, passenger car with internal   combustion engine | Cut-off, U",
    "Transport, passenger car, electric, with Swiss electricity mix {GLO}| transport, passenger car, electric | Cut-off, U",
    "Transport, passenger, motor scooter {CH}| transport, passenger, motor scooter | Cut-off, U",
    "Transport, passenger, bicycle {CH}| transport, passenger, bicycle | Cut-off, U",
    "Transport, regular bus {CH}| transport, regular bus | Cut-off, U",
    "Transport, tram {CH}| transport, tram | Cut-off, U",
    "Transport, freight, sea, ferry {GLO}| transport, freight, sea, ferry | Cut-off, U"
  )
)

kable(transport_data)

```

<br><br>

### IPCC GWP 100a Impact Assessment Method

To translate life cycle inventory results into climate-relevant metrics, the IPCC 2021 GWP 100a impact assessment method was applied [@ipcc2021gwp100]. This method expresses the global warming potential (GWP) of greenhouse gases in terms of carbon dioxide equivalents (CO₂e) over a 100-year time horizon. It reflects the most recent scientific consensus provided by the Intergovernmental Panel on Climate Change (IPCC) and enables standardized comparison of different emission sources and activities. By applying this method within SimaPro, all modeled greenhouse gas emissions were converted into consistent and comparable CO₂e values, ensuring the results align with international reporting standards and climate targets.

<br><br>

# Results 

This section presents key findings from our two-month mobility analysis. We first examine the distribution between rest and movement, followed by an overview of the most visited locations. We then analyse individual transport activities, including travel distances and CO₂ emissions by mode, and conclude with a comparison to the Swiss average.

```{r, warning=FALSE, message=FALSE}
start_datum <- ymd_hms("2025-02-06 11:40:35")
end_datum   <- ymd_hms("2025-04-06 23:59:00")
dauer_gesamt <- as.numeric(difftime(end_datum, start_datum, units = "hours"))
```

```{r}
#| label: load-data
#| cache: true

file_dimi <- "data/timeline_dimi_new.json"
data_dimi <- fromJSON(file_dimi)

file_ramon <- "data/Timeline_ramon_new.json"
data_ramon <- fromJSON(file_ramon)
```

```{r}
#| eval: false

# activity_df_dimi <- data_dimi$semanticSegments$activity %>%
#   mutate(
#     start_latlng = map_chr(start$latLng, ~ .x %||% NA_character_),
#     end_latlng = map_chr(end$latLng, ~ .x %||% NA_character_),
#     distance = distanceMeters,
#     activity_probability = probability,
#     type = topCandidate$type,
#     type_probability = topCandidate$probability,
#     parking_latlng = map_chr(parking$location$latLng, ~ .x %||% NA_character_)
#   ) %>%
#   select(start_latlng, end_latlng, distance, activity_probability, type, type_probability, parking_latlng)
# 
# activity_clean_dimi <- data_dimi$semanticSegments$activity %>%
#   filter(
#     !is.na(start$latLng) |
#     !is.na(end$latLng) |
#     !is.na(distanceMeters) |
#     !is.na(topCandidate$type) |
#     !is.na(topCandidate$probability)
#   )
# 

```

```{r}
#| eval: false

# activity_df_ramon <- data_ramon$semanticSegments$activity %>%
#   mutate(
#     start_latlng = map_chr(start$latLng, ~ .x %||% NA_character_),
#     end_latlng = map_chr(end$latLng, ~ .x %||% NA_character_),
#     distance = distanceMeters,
#     activity_probability = probability,
#     type = topCandidate$type,
#     type_probability = topCandidate$probability,
#     parking_latlng = map_chr(parking$location$latLng, ~ .x %||% NA_character_)
#   ) %>%
#   select(start_latlng, end_latlng, distance, activity_probability, type, type_probability, parking_latlng)
# 
# activity_clean_ramon <- data_ramon$semanticSegments$activity %>%
#   filter(
#     !is.na(start$latLng) |
#     !is.na(end$latLng) |
#     !is.na(distanceMeters) |
#     !is.na(topCandidate$type) |
#     !is.na(topCandidate$probability)
#   )


```

```{r}
#| eval: false

# # Nur gültige timelinePath-Elemente
# valid_indices <- which(map_lgl(data_ramon$semanticSegments$timelinePath, ~ is.data.frame(.x)))
# 
# # DataFrame mit Start-/Endzeit + Punktdaten
# paths_df_clean_ramon <- map_dfr(valid_indices, function(i) {
#   path <- data_ramon$semanticSegments$timelinePath[[i]]
#   start_time <- data_ramon$semanticSegments$startTime[i]
#   end_time   <- data_ramon$semanticSegments$endTime[i]
#   
#   tibble(
#     segment_id    = i,
#     point         = path$point,
#     time          = ymd_hms(str_sub(path$time, 1, 19)),  # Zeit säubern
#     segment_start = ymd_hms(start_time),
#     segment_end   = ymd_hms(end_time)
#   )
# }) %>%
#   separate(point, into = c("lat", "lng"), sep = ", ") %>%
#   mutate(
#     lat = str_remove(lat, "°") %>% as.numeric(),
#     lng = str_remove(lng, "°") %>% as.numeric()
#   )

```

```{r}
#| eval: false

# # Nur gültige timelinePath-Elemente
# valid_indices <- which(map_lgl(data_dimi$semanticSegments$timelinePath, ~ is.data.frame(.x)))
# 
# # DataFrame mit Start-/Endzeit + Punktdaten
# paths_df_clean_dimi <- map_dfr(valid_indices, function(i) {
#   path <- data_dimi$semanticSegments$timelinePath[[i]]
#   start_time <- data_dimi$semanticSegments$startTime[i]
#   end_time   <- data_dimi$semanticSegments$endTime[i]
#   
#   tibble(
#     segment_id    = i,
#     point         = path$point,
#     time          = ymd_hms(str_sub(path$time, 1, 19)),  # Zeit säubern
#     segment_start = ymd_hms(start_time),
#     segment_end   = ymd_hms(end_time)
#   )
# }) %>%
#   separate(point, into = c("lat", "lng"), sep = ", ") %>%
#   mutate(
#     lat = str_remove(lat, "°") %>% as.numeric(),
#     lng = str_remove(lng, "°") %>% as.numeric()
#   )
# 

```



```{r}
#| warning: false

#Start visits

visits_dimi <- data_dimi$semanticSegments |> 
  filter(!is.na(visit$topCandidate$placeLocation$latLng)) |> 
  mutate(
    place_id = visit$topCandidate$placeId,
    lat_lng = visit$topCandidate$placeLocation$latLng,
    probability = visit$probability,
    start = ymd_hms(startTime),
    end = ymd_hms(endTime)
  ) %>%
  select(start, end, place_id, lat_lng, probability)



# Dimi: Filter visits longer than 60 min and split coordinates
visits_dimi_2 <- data_dimi$semanticSegments |> 
  filter(!is.na(visit$topCandidate$placeLocation$latLng)) |> 
  mutate(
    place_id = visit$topCandidate$placeId,
    lat_lng = visit$topCandidate$placeLocation$latLng,
    start = ymd_hms(startTime),
    end = ymd_hms(endTime),
    duration_mins = as.numeric(difftime(end, start, units = "mins"))
  ) |> 
  separate(lat_lng, into = c("lat", "lng"), sep = ", ") |> 
  mutate(
    lat = str_remove(lat, "°") %>% as.numeric(),
    lng = str_remove(lng, "°") %>% as.numeric()
  ) |> 
  filter(duration_mins >= 60) |> 
  select(start, end, duration_mins, lat, lng, place_id)

# Dimi: Leaflet Map
leaflet_map_dimi <- leaflet(visits_dimi_2) |> 
  addTiles() %>%
  addCircleMarkers(
    ~lng, ~lat,
    radius = 3,
    color = "blue",
    stroke = FALSE,
    fillOpacity = 0.7,
    popup = ~paste0(
      "<b>Start:</b> ", start, "<br>",
      "<b>Ende:</b> ", end, "<br>",
      "<b>Dauer:</b> ", round(duration_mins), " Minuten"
    )
  ) %>%
  addLegend(
    position = "bottomright",
    colors = "blue",
    labels = "Dimi's Visits > 1 hour",
    title = "Google Timeline"
  )
```


```{r}
#| eval: false

# Speichere die HTML-Version
saveWidget(leaflet_map_dimi, "figures/tmp_map.html", selfcontained = TRUE)

# Screenshot als PNG speichern
webshot("figures/tmp_map.html", file = "figures/dimi_visits_map.png", vwidth = 800, vheight = 600)
```


```{r}
#| warning: false

# Ramon: Alle Aufenthalte extrahieren
visits_ramon <- data_ramon$semanticSegments |> 
  filter(!is.na(visit$topCandidate$placeLocation$latLng)) |> 
  mutate(
    place_id = visit$topCandidate$placeId,
    lat_lng = visit$topCandidate$placeLocation$latLng,
    probability = visit$probability,
    start = ymd_hms(startTime),
    end = ymd_hms(endTime)
  ) %>%
  select(start, end, place_id, lat_lng, probability)

visits_ramon_2 <- data_ramon$semanticSegments |> 
  filter(!is.na(visit$topCandidate$placeLocation$latLng)) |> 
  mutate(
    place_id = visit$topCandidate$placeId,
    lat_lng = visit$topCandidate$placeLocation$latLng,
    start = ymd_hms(startTime),
    end = ymd_hms(endTime),
    duration_mins = as.numeric(difftime(end, start, units = "mins"))
  ) |> 
  separate(lat_lng, into = c("lat", "lng"), sep = ", ") |> 
  mutate(
    lat = str_remove(lat, "°") %>% as.numeric(),
    lng = str_remove(lng, "°") %>% as.numeric()
  ) |> 
  filter(duration_mins >= 60) |> 
  select(start, end, duration_mins, lat, lng, place_id)

leaflet_map_ramon <- leaflet(visits_ramon_2) |> 
  addTiles() %>%
  addCircleMarkers(
    ~lng, ~lat,
    radius = 3,
    color = "red",
    stroke = FALSE,
    fillOpacity = 0.7,
    popup = ~paste0(
      "<b>Start:</b> ", start, "<br>",
      "<b>Ende:</b> ", end, "<br>",
      "<b>Dauer:</b> ", round(duration_mins), " Minuten"
    )
  ) %>%
  addLegend(
    position = "bottomright",
    colors = "red",
    labels = "Ramon's visits > 1 Stunde",
    title = "Google Timeline"
  )
```


```{r}
#| eval: false

# Speichere die HTML-Version
saveWidget(leaflet_map_ramon, "figures/tmp_map.html", selfcontained = TRUE)

# Screenshot als PNG speichern
webshot("figures/tmp_map.html", file = "figures/ramon_visits_map.png", vwidth = 800, vheight = 600)
```

```{r}
#| echo: false

## Visits

#Ramon

# Visits filtern nach Zeitraum
visits_ramon_filtered <- visits_ramon_2 %>%
  filter(start >= start_datum & end <= end_datum)

visits_min_ramon <- sum(visits_ramon_filtered$duration_mins)

visits_h_ramon <- visits_min_ramon/60

visits_day_ramon <- visits_h_ramon/24

```

```{r}
#Dimi
# visits
visits_dimi_filtered <- visits_dimi_2 %>%
  filter(start >= start_datum & end <= end_datum)

visits_min_dimi <- sum(visits_dimi_filtered$duration_mins)

visits_h_dimi <- visits_min_dimi/60

visits_day_dimi <- visits_h_dimi/24

```

```{r}
#| echo: false

## Activity

#dimi
segments_dimi <- data_dimi$semanticSegments

# Extrahiere und filtere Dimi's Aktivitätsdaten
activity_dimi <- segments_dimi |> 
  filter(!is.na(activity$topCandidate$type)) |> 
  mutate(
    activity_type = ifelse(activity$topCandidate$type == "IN_FERRY", "SKIING", activity$topCandidate$type),
    probability = activity$topCandidate$probability,
    distance_km = round(activity$distanceMeters / 1000, 2),
    start_latlng = activity$start$latLng,
    end_latlng = activity$end$latLng,
    start = ymd_hms(startTime),
    end = ymd_hms(endTime),
    duration_mins = as.numeric(difftime(end, start, units = "mins")),
    speed_kmh = round((distance_km) / (duration_mins / 60), 2)
  ) |> 
  filter(start >= start_datum, end <= end_datum) |> 
  select(start, end, duration_mins, distance_km, speed_kmh,
         start_latlng, end_latlng, activity_type, probability)


activity_min_dimi<- sum(activity_dimi$duration_mins)

activity_h_dimi<- activity_min_dimi/60

act_visit_dimi <- activity_h_dimi + visits_h_dimi

activity_day_dimi <- activity_h_dimi/24


# Dauer berechnen
dauer_gesamt_dimi_h <- difftime(end_datum, start_datum, units = "hours")


dauer_gesamt_dimi_day <- difftime(end_datum, start_datum, units = "day")



nan_h_dimi <- as.numeric(dauer_gesamt_dimi_h) - (activity_h_dimi + visits_h_dimi)

nan_day_dimi <- as.numeric(dauer_gesamt_dimi_day) - (activity_day_dimi + visits_day_dimi)



summe_stunden_dimi <- activity_h_dimi + visits_h_dimi + nan_h_dimi

```


```{r}
#ramon
segments_ramon <- data_ramon$semanticSegments

activity_ramon <- segments_ramon |> 
  filter(!is.na(activity$topCandidate$type)) |> 
  mutate(
    activity_type = ifelse(activity$topCandidate$type == "IN_SUBWAY", "IN_TRAIN", activity$topCandidate$type),
    probability = activity$topCandidate$probability,
    distance_km = round(activity$distanceMeters / 1000, 2),
    start_latlng = activity$start$latLng,
    end_latlng = activity$end$latLng,
    start = ymd_hms(startTime),
    end = ymd_hms(endTime),
    duration_mins = as.numeric(difftime(end, start, units = "mins")),
    speed_kmh = round((distance_km) / (duration_mins / 60), 2)
  ) |> 
  filter(start >= start_datum, end <= end_datum) |> 
  select(start, end, duration_mins, distance_km, speed_kmh,
         start_latlng, end_latlng, activity_type, probability)

activity_min_ramon<- sum(activity_ramon$duration_mins)

activity_h_ramon<- activity_min_ramon/60

act_visit_ramon <- activity_h_ramon + visits_h_ramon 

activity_day_ramon <- activity_h_ramon/24



# Dauer in Stunden berechnen
dauer_gesamt_ramon <- difftime(end_datum, start_datum, units = "hours")


dauer_gesamt_ramon_day <- difftime(end_datum, start_datum, units = "day")



nan_h_ramon <- as.numeric(dauer_gesamt_ramon) - (activity_h_ramon + visits_h_ramon)

nan_day_ramon <- as.numeric(dauer_gesamt_ramon_day) - (activity_day_ramon + visits_day_ramon)

summe_stunden_ramon <- activity_h_ramon + visits_h_ramon + nan_h_ramon
```

<br><br>

## Distribution between rest and movement {#sec-Distribution}
As illustrated in @fig-rest-movement, the majority of the Google Timeline data is classified as visits (e.g. staying at home, working), while activities (such as walking, cycling or driving) account for a much smaller share. When aggregating the total recorded time, Dimi’s activities correspond to 6.5 full days, and Ramon’s to 6 days. The remaining recorded time – 51.5 and 52 days respectively – was spent at stationary locations. This highlights that although movement occurred regularly, it made up only a small fraction of the total recorded time.
```{r}
#ramon
round_half <- function(x) round(x * 2) / 2


# DataFrame
bewegung_df_ramon <- tibble(
  status = c("Activity", "Visits", "No Data"),
  stunden = c(activity_h_ramon, visits_h_ramon, nan_h_ramon)
) %>%
  mutate(
    label = paste0(status, "\n", round(stunden), " h"),
    ypos = cumsum(stunden) - 0.5 * stunden
  )

#für tage
bewegung_df_ramon_days <- tibble(
  status = c("Activity", "Visits", "No Data"),
  tage = c(activity_h_ramon, visits_h_ramon, nan_h_ramon) / 24
) %>%
  mutate(
    tage = round_half(tage),
    status_label = paste0(status, " (", tage, "d)")
  )




bewegung_df_ramon <- bewegung_df_ramon %>%
  mutate(status_label = paste0(status, " (", round(stunden), "h)"))
```


```{r}
#| eval: false


ggplot(bewegung_df_ramon, aes(x = "", y = stunden, fill = status_label)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  labs(
    title = "Share of activities, visits and no data in hours (Ramon)",
    fill = "Status",
    y = NULL, x = NULL
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11)
  )
```


```{r}

# Plot
pie_plot_ramon_days <- ggplot(bewegung_df_ramon_days, aes(x = "", y = tage, fill = status_label)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  labs(
    fill = "Status",
    y = NULL, x = NULL
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11)
  )

```


```{r}
#| eval: false

ggsave("figures/pie_plot_ramon_days.png", pie_plot_ramon_days, width = 6, height = 6)


```



```{r}

# DataFrame
bewegung_df_dimi <- tibble(
  status = c("Activity", "Visits", "No Data"),
  stunden = c(activity_h_dimi, visits_h_dimi, nan_h_dimi)
) %>%
  mutate(
    label = paste0(status, "\n", round(stunden), " h"),
    ypos = cumsum(stunden) - 0.5 * stunden
  )

bewegung_df_dimi_days <- tibble(
  status = c("Activity", "Visits", "No Data"),
  tage = c(activity_h_dimi, visits_h_dimi, nan_h_dimi) / 24
) %>%
  mutate(
    tage = round_half(tage),
    status_label = paste0(status, " (", tage, "d)")
  )


bewegung_df_dimi <- bewegung_df_dimi %>%
  mutate(status_label = paste0(status, " (", round(stunden), "h)"))
```


```{r}
#| eval: false

ggplot(bewegung_df_dimi, aes(x = "", y = stunden, fill = status_label)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  labs(
    title = "Share of activities, visits and no data in hours (Dimi)",
    fill = "Status",
    y = NULL, x = NULL
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11)
  )
```


```{r}

# Plot tage
pie_plot_dimi_days <- ggplot(bewegung_df_dimi_days, aes(x = "", y = tage, fill = status_label)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  labs(
    fill = "Status",
    y = NULL, x = NULL
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11)
  )


```


```{r}
#| eval: false

ggsave("figures/pie_plot_dimi_days.png", pie_plot_dimi_days, width = 6, height = 6)

```


```{r}
pie_plot_dimi_days_labeled <- pie_plot_dimi_days +
  labs(subtitle = "(a) Dimi") +
  theme(plot.subtitle = element_text(hjust = 0.5, size = 12))

pie_plot_ramon_days_labeled <- pie_plot_ramon_days +
  labs(subtitle = "(b) Ramon") +
  theme(plot.subtitle = element_text(hjust = 0.5, size = 12))


# Kombinierte Abbildung
combined_plot <- pie_plot_dimi_days_labeled + pie_plot_ramon_days_labeled +
  plot_layout(ncol = 2) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14)
  )
```

```{r}
#| label: fig-rest-movement
#| fig-cap: "Distribution of activities, visits and no data in days"

combined_plot

```

<br><br>

## Visits

As shown in @fig-map-dimi, Dimi visited several locations within Switzerland and also spent time in Spain. Notably, visits to Madrid and Barcelona were recorded, both of which were related to work. The spatial distribution is therefore more dispersed, reflecting international travel activity in addition to domestic mobility. The geographic data were visualized using OpenStreetMap [@openstreetmap2025].


```{r}
#| label: fig-map-dimi
#| fig-cap: "Visited locations of Dimi [@openstreetmap2025]"

leaflet_map_dimi
```

<br><br>

@fig-map-ramon shows that Ramon's mobility was exclusively within Switzerland. The data reveals a dense pattern of visits in central Switzerland and parts of western Switzerland, indicating a regionally focused mobility behaviour. Most of the longer trips were related to ice hockey games that Ramon played during the observed period.
```{r}
#| label: fig-map-ramon
#| fig-cap: "Visited locations of Ramon [@openstreetmap2025]"
#| 
leaflet_map_ramon
```

<br><br>

## Activities
```{r}
#| warning: false


co2_data_dimi <- read_delim("data/transport_co2_factors.csv")

co2_data_ramon <- read_delim("data/transport_co2_factors.csv")

co2_table <- tribble(
  ~activity_type,             ~co2_kg_per_km,
  "IN_PASSENGER_VEHICLE",     0.111,
  "WALKING",                  0.000,
  "IN_BUS",                   0.114,
  "IN_TRAIN",                 0.011,
  "IN_TRAM",                  0.016,
  "CYCLING",                  0.012,
  "IN_FERRY",                 0.009,
  "IN_SUBWAY",                0.016,
  "MOTORCYCLING",             0.121,
  "SKIING",                   0.000
) %>%
  mutate(activity_label = recode(
    activity_type,
    "CYCLING" = "Cycling",
    "IN_BUS" = "Bus",
    "IN_FERRY" = "Ferry",
    "IN_PASSENGER_VEHICLE" = "Car",
    "IN_SUBWAY" = "Subway",
    "IN_TRAIN" = "Train",
    "IN_TRAM" = "Tram",
    "MOTORCYCLING" = "Motorcycle",
    "SKIING" = "Skiing",
    "WALKING" = "Walking"
  )) %>%
  select(activity_label, co2_kg_per_km)

#knitr::kable(co2_table, caption = "CO₂-Faktoren nach Transportmittel")
```




```{r}

#co2_table %>%
 # gt() %>%
  #tab_header(title = md("**CO₂ eq-factors per transportmode**")) %>%
#  cols_label(
#    activity_label = "Transportmode",
#    co2_kg_per_km = "CO₂ eq in kg per passenger kilometer"
#  ) %>%
#  fmt_number(columns = co2_kg_per_km, decimals = 3)

activity_dimi_co2 <- full_join(activity_dimi, co2_data_dimi, by = "activity_type") |> 
  mutate(
    co2_emission_in_kg = round(distance_km * co2_kg_per_km, 2)
  ) |> 
  select(
    start, end, duration_mins, distance_km, speed_kmh,
    start_latlng, end_latlng, activity_type, probability,
    co2_kg_per_km, co2_emission_in_kg
  )

activity_ramon_co2 <- full_join(activity_ramon, co2_data_ramon, by = "activity_type") |> 
  mutate(
    co2_emission_in_kg = round(distance_km * co2_kg_per_km, 2)
  ) |> 
  select(
    start, end, duration_mins, distance_km, speed_kmh,
    start_latlng, end_latlng, activity_type, probability,
    co2_kg_per_km, co2_emission_in_kg
  )

```

```{r}
# 📊 Summarize CO₂, distance, time (in hours), speed, and emissions per hour for Dimi
emissions_summary_dimi <- activity_dimi_co2 |> 
  group_by(activity_type) |> 
  summarise(
    total_distance_km = round(sum(distance_km, na.rm = TRUE), 2),
    total_duration_hrs = round(sum(duration_mins, na.rm = TRUE) / 60, 2),
    total_emission_kg = round(sum(co2_emission_in_kg, na.rm = TRUE), 2),
    average_speed_kmh = round(total_distance_km / total_duration_hrs, 2),
    co2_kg_per_hour = round(total_emission_kg / total_duration_hrs, 2),
    co2_kg_per_km = round(total_emission_kg / total_distance_km, 2),
    .groups = "drop"
  )

# ➕ Add TOTAL row
emissions_total_dimi <- emissions_summary_dimi |> 
  summarise(
    activity_type = "TOTAL",
    total_distance_km = sum(total_distance_km),
    total_duration_hrs = sum(total_duration_hrs),
    total_emission_kg = sum(total_emission_kg),
    average_speed_kmh = round(total_distance_km / total_duration_hrs, 2),
    co2_kg_per_hour = round(total_emission_kg / total_duration_hrs, 2),
    co2_kg_per_km = round(total_emission_kg / total_distance_km, 2),

  )

# 📦 Combine
emissions_summary_full_dimi <- bind_rows(emissions_summary_dimi, emissions_total_dimi)

emissions_summary_full_dimi <- emissions_summary_full_dimi %>%
  filter(!activity_type %in% c("MOTORCYCLING", "IN_FERRY"))

# 👀 View result
#print(emissions_summary_full_dimi)
```

```{r}
# 📊 Summarize CO₂, distance, time (in hours), speed, and emissions per hour for Ramon
emissions_summary_ramon <- activity_ramon_co2 |> 
  group_by(activity_type) |> 
  summarise(
    total_distance_km = round(sum(distance_km, na.rm = TRUE), 2),
    total_duration_hrs = round(sum(duration_mins, na.rm = TRUE) / 60, 2),
    total_emission_kg = round(sum(co2_emission_in_kg, na.rm = TRUE), 2),
    average_speed_kmh = round(total_distance_km / total_duration_hrs, 2),
    co2_kg_per_hour = round(total_emission_kg / total_duration_hrs, 2),
    co2_kg_per_km = round(total_emission_kg / total_distance_km, 2),
    .groups = "drop"
  )

# ➕ Add TOTAL row
emissions_total_ramon <- emissions_summary_ramon |> 
  summarise(
    activity_type = "TOTAL",
    total_distance_km = sum(total_distance_km),
    total_duration_hrs = sum(total_duration_hrs),
    total_emission_kg = sum(total_emission_kg),
    average_speed_kmh = round(total_distance_km / total_duration_hrs, 2),
    co2_kg_per_hour = round(total_emission_kg / total_duration_hrs, 2),
    co2_kg_per_km = round(total_emission_kg / total_distance_km, 2)
  )

# 📦 Combine
emissions_summary_full_ramon <- bind_rows(emissions_summary_ramon, emissions_total_ramon)

emissions_summary_full_ramon <- emissions_summary_full_ramon %>%
  filter(!activity_type %in% c("IN_SUBWAY"))

# 👀 View result
#print(emissions_summary_full_ramon)
```


```{r}
# 📦 Add person column to each dataset
emissions_dimi_labeled <- emissions_summary_full_dimi |> 
  mutate(person = "Dimi")

emissions_ramon_labeled <- emissions_summary_full_ramon |> 
  mutate(person = "Ramon")

# 🧱 Combine both
emissions_combined <- bind_rows(emissions_dimi_labeled, emissions_ramon_labeled)

# 🚶‍♂️ Personen und Transportmodi definieren
personen <- c("Dimi", "Ramon")
aktivitaeten <- c(
  "Subway", "Tram", "Motorcycle", "Walking", "Skiing", "Ferry", 
  "Cycling", "Bus", "Car", "Train"
)

# 🧼 Label-Mapping anwenden
label_mapping <- c(
  "CYCLING" = "Cycling",
  "IN_BUS" = "Bus",
  "IN_FERRY" = "Ferry",
  "IN_PASSENGER_VEHICLE" = "Car",
  "IN_SUBWAY" = "Subway",
  "IN_TRAIN" = "Train",
  "IN_TRAM" = "Tram",
  "MOTORCYCLING" = "Motorcycle",
  "SKIING" = "Skiing",
  "WALKING" = "Walking",
  "TOTAL" = "Total"
)

# Nur relevante Aktivitäten und Labels zuweisen
emissions_combined_filtered <- emissions_combined |> 
  filter(activity_type != "TOTAL") |> 
  mutate(activity_label = label_mapping[activity_type])

# 🔄 Alle Kombinationen von Personen und Aktivitäten
kombis <- expand.grid(
  person = personen,
  activity_label = aktivitaeten,
  stringsAsFactors = FALSE
)
kombis$activity_label <- factor(kombis$activity_label, levels = aktivitaeten)

# 🔁 Join mit kombis → fehlende Werte werden ergänzt
emissions_padded <- kombis |> 
  left_join(emissions_combined_filtered, by = c("person", "activity_label")) |> 
  mutate(
    activity_label = factor(activity_label, levels = aktivitaeten),
    total_emission_kg = ifelse(total_emission_kg < 1, NA, total_emission_kg)
  )

# 🔢 Reihenfolge für den x-Achsen-Faktor festlegen
emissions_padded$activity_label <- factor(
  emissions_padded$activity_label,
  levels = c(
    "Subway", "Tram", "Walking", "Skiing", "Ferry", 
    "Motorcycle", "Cycling", "Train", "Bus", "Car"
  )
)


# 📊 Plot 1: Emissionen
emissions_plot <- ggplot(emissions_padded, aes(x = activity_label, y = total_emission_kg, fill = person)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), na.rm = TRUE) +
  scale_y_log10(
    limits = c(1, 200),
    breaks = c(1, 10, 50, 100, 200),
    labels = scales::label_number(big.mark = "'")
  ) +
  labs(
    title = expression("CO"[2] * " Emissions by Transport Mode"),
    x = "Transport Mode",
    y = expression("CO"[2] * " Emissions in kg (log)")
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title = element_text(size = 14),
    plot.title = element_text(size = 16, hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13)
  )

# 🔢 Reihenfolge für den x-Achsen-Faktor festlegen
emissions_padded$activity_label <- factor(
  emissions_padded$activity_label,
  levels = c(
    "Subway", "Tram", "Motorcycle", "Walking", "Skiing", "Ferry", 
  "Cycling", "Bus", "Car", "Train"
  )
)

# 📊 Plot 2: Distanz
distance_plot <- ggplot(emissions_padded, aes(x = activity_label, y = total_distance_km, fill = person)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), na.rm = TRUE) +
  scale_y_log10(
    limits = c(1, 10000),
    breaks = c(1, 10, 100, 1000, 10000),
    labels = scales::label_number(big.mark = "'")
  ) +
  labs(
    title = "Distance in km by Activity Type",
    x = "Transport Mode",
    y = "Distance in km (log)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title = element_text(size = 14),
    plot.title = element_text(size = 16, hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13)
  )

```


@fig-Total-Distance displays the total distances in kilometers travelled by Dimi and Ramon, broken down by transport mode. Each bar represents the overall distance per person and is subdivided by transport modes.

Dimi travelled a total of 6,878 kilometers, more than double Ramon's total of 2,730 kilometers.

For Dimi, the dominant transport mode is Train, accounting for approximately 4,938 km, followed by Car with 1,361 km, and Bus with 197 km. Other modes such as Cycling, Walking, Tram, Subway, Skiing, and Ferry contribute smaller portions to the total.
A significant portion of Dimi’s train distance — roughly 3,100 km — is due to a single long-distance trip to Spain. Remarkably, this single journey alone exceeds Ramon’s entire travel distance over the two-month period.

Ramon’s top transport mode is Bus, with 1,188 km, followed by Train (975 km) and Car (305 km). Notably, Motorcycling appears in Ramon's data (24 km) but is absent for Dimi. In contrast, Ramon has no recorded distance for Skiing or Subway, both of which are present in Dimi’s profile.

This visualization highlights not only the difference in total mobility between the two individuals but also their distinct preferences in transport modes.


```{r}
#| label: fig-Total-Distance
#| fig-cap: Total Distance travelled by Person and Transport Mode


set.seed(123)  # für reproduzierbare Farben

# 1. Alle Labels erfassen
all_modes <- sort(unique(emissions_combined_filtered$activity_label))
n_modes <- length(all_modes)

# 2. Farben generieren (zufällig aus Set3 + zusätzlichen Farben)
base_colors <- c(brewer.pal(12, "Set3"),
                 "#999999", "#CCCCCC", "#666666", "#FFB6C1", "#87CEFA", "#8dd3c7", "#fb8072")

# 3. Farben mischen und zuweisen
shuffled_colors <- sample(base_colors, size = n_modes)
transport_colors <- setNames(shuffled_colors, all_modes)

# Reihenfolge der Transportmittel (optional visuell)
dimi_sorted_km <- emissions_combined_filtered %>%
  group_by(activity_label) %>%
  summarise(total_km = sum(total_distance_km), .groups = "drop") %>%
  arrange(total_km) %>%
  pull(activity_label)

# Daten vorbereiten
plot_data_km <- emissions_combined_filtered %>%
  group_by(person, activity_label) %>%
  summarise(total_km = sum(total_distance_km), .groups = "drop") %>%
  mutate(activity_label = factor(activity_label, levels = dimi_sorted_km))

# Totalsummen
totals_km <- plot_data_km %>%
  group_by(person) %>%
  summarise(total = sum(total_km))

text_offset_km <- 0.05 * max(totals_km$total)

# Plot
ggplot(plot_data_km, aes(x = person, y = total_km, fill = activity_label)) +
  geom_bar(stat = "identity") +
  geom_text(
    data = totals_km,
    aes(x = person, y = total + text_offset_km, label = paste0(round(total), " km")),
    inherit.aes = FALSE,
    size = 5
  ) +
  labs(
    x = "Person",
    y = "Distance (km)",
    fill = "Transport Mode"
  ) +
  scale_fill_manual(values = transport_colors, na.translate = FALSE) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title = element_text(size = 14),
    plot.title = element_text(hjust = 0.5, size = 16),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13)
  )

```

<br><br>

As shown in @tbl-co2-factors, different transport modes vary significantly in their CO₂ equivalent emissions per passenger kilometer. Surprisingly, cycling does not have a CO₂e value of zero. This is because the value accounts for life-cycle emissions, including those from the production and maintenance of the bicycle. Since these emissions are allocated per passenger kilometer, cycling ends up having a slightly higher emission factor than train travel. For skiing, no specific CO₂ equivalent emission factor could be identified, so we assumed it to be zero.

For the ferry trip, we considered that only Ramon used this mode of transport. Although it was a freight ferry, we estimated the emissions based on Ramon’s body weight to approximate a fair share of per-person emissions in the absence of more granular data.

In the case of car travel, we assumed an average occupancy of 1.5 persons per vehicle, consistent with standard assumptions in emission modelling. Furthermore, we estimated that 75% of the distance travelled by car was covered using an electric vehicle (EV), while the remaining 25% was driven with a conventional internal combustion engine (ICE) vehicle. Accordingly, the total car emissions were calculated as a weighted average: 75% from EV emission factors and 25% from conventional vehicle factors. 
```{r}
#| label: tbl-co2-factors
#| tbl-cap: CO₂ eq-factors per transportmode according to @ipcc2021gwp100.

co2_table %>%
  kable(
    col.names = c("Transportmode", "CO₂eq in kg per passenger kilometer"),
    digits = 3,
    align = "l"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE
  )

```

<br><br>

@fig-co2-km-bubble provides a comprehensive overview of both mobility behavior and resulting CO₂e emissions for Dimi and Ramon across various transport modes. The y-axis represents the total distance travelled (log scale), while the size of each bubble corresponds to the total CO₂e emissions (in kilograms) for that mode.

This visualisation highlights the differences in transport choices and their environmental impacts. While Dimi travelled further overall – covering around 5,000 km by train and 1,300 km by car – his car-related emissions were still three times higher than those from train travel. This is particularly noteworthy given that approximately 75% of his car travel was done using an electric vehicle. The result underlines the high efficiency of rail transport, even when compared to low-emission vehicle alternatives such as electric cars.

Ramon, on the other hand, showed a more regionally focused travel pattern but still generated substantial emissions from bus usage. The contrast between high distances and small bubbles for train travel, versus relatively large bubbles for car and bus use, clearly illustrates the importance of low-emission transport options in reducing one’s carbon footprint.


```{r}
#| fig-align: center
#| label: fig-co2-km-bubble
#| fig-cap: CO₂e emissions (bubble size) and distance travelled (log-scale) per transport mode for Dimi and Ramon

#final plot



emissions_combined <- emissions_combined %>%
  mutate(activity_label = recode(activity_type, !!!label_mapping))

emissions_combined$activity_label <- factor(
  emissions_combined$activity_label,
  levels = c(
    "Subway", "Tram", "Motorcycle", "Walking", "Skiing", "Ferry", 
    "Cycling", "Bus", "Car", "Train", "Total"
  )
)


final_plot <- ggplot(emissions_combined, aes(
  x = activity_label,
  y = total_distance_km,
  size = total_emission_kg,
  color = person
)) +
  geom_point(alpha = 0.8) +
  geom_text(
  data = emissions_combined %>% filter(total_emission_kg >= 0),
  mapping = aes(
    x = activity_label,
    y = total_distance_km,
    label = ifelse(
      total_emission_kg < 1,
      paste0(round(total_emission_kg, 1), " kg"),
      paste0(round(total_emission_kg), " kg")
    ),
    vjust = ifelse(total_emission_kg >= 100, -1.2, -0.8),
    hjust = ifelse(total_emission_kg >= 100, 1.5, 1.4)
  ),
  inherit.aes = FALSE,
  size = 4,
  show.legend = FALSE
) +
scale_y_log10(
  breaks = c(1, 10, 100, 1000, 10000),
  labels = scales::label_number(big.mark = "'"), 
  limits = c(5, 10000)
) +  
  scale_size_continuous(
    range = c(4, 20),
    breaks = seq(0, max(emissions_combined$total_emission_kg, na.rm = TRUE), by = 50),
    limits = c(0, max(emissions_combined$total_emission_kg, na.rm = TRUE))
  ) +
  labs(
    #title = expression("CO"[2] * " Emissions by Transport Mode"),
    x = "Transport Mode",
    y = "Distance Traveled in km (log)",
    size = expression("CO"[2] * " Emissions (kg)"),
    color = "Person"
  ) +
  scale_color_manual(
    values = c("Dimi" = "#FF6F61", "Ramon" = "#00BFC4"),
    breaks = c("Dimi", "Ramon")
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title = element_text(size = 14),
    plot.title = element_text(hjust = 0.5, size = 16),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13)
  )

final_plot


ggsave("figures/final_plot.png", final_plot, width = 6, height = 6)

```


```{r}
#| eval: false

# 1. Emissionen zusammenfassen
df_emissions <- emissions_combined_filtered %>%
  group_by(person, activity_label) %>%
  summarise(value = sum(total_emission_kg), .groups = "drop") %>%
  mutate(type = "Emissions")

# 2. Distanzen zusammenfassen
df_distance <- emissions_combined_filtered %>%
  group_by(person, activity_label) %>%
  summarise(value = sum(total_distance_km), .groups = "drop") %>%
  mutate(type = "Distance")

# 3. Kombinieren
df_combined <- bind_rows(df_emissions, df_distance)

# 4. Totals berechnen für Verhältnis
totals <- df_combined %>%
  group_by(person, type) %>%
  summarise(total = sum(value), .groups = "drop")

em_dimi <- totals %>% filter(person == "Dimi", type == "Emissions") %>% pull(total)
em_ramon <- totals %>% filter(person == "Ramon", type == "Emissions") %>% pull(total)
km_dimi <- totals %>% filter(person == "Dimi", type == "Distance") %>% pull(total)
km_ramon <- totals %>% filter(person == "Ramon", type == "Distance") %>% pull(total)

# Verhältnis (Wurzel für Fläche = visuell sinnvoller)
em_ratio <- sqrt(em_ramon / em_dimi)
km_ratio <- sqrt(km_ramon / km_dimi)

# 5. Plotfunktion (beschriftet, keine Legende)
plot_pie_noleg <- function(pie_data, person_name, value_type, label_threshold = 0.05) {
  df <- pie_data %>%
    filter(person == person_name, type == value_type) %>%
    mutate(
      share = value / sum(value),
      label = ifelse(share > label_threshold, activity_label, "")
    )

  ggplot(df, aes(x = "", y = value, fill = activity_label)) +
    geom_col(width = 1, color = "white") +
    coord_polar(theta = "y") +
    geom_text(aes(label = label), position = position_stack(vjust = 0.5), size = 3) +
    labs(title = paste(person_name, "-", value_type)) +
    theme_void() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 14),
      legend.position = "none"
    )
}

# 6. Plots generieren
p1 <- plot_pie_noleg(df_combined, "Dimi", "Emissions")
p2 <- plot_pie_noleg(df_combined, "Ramon", "Emissions")
p3 <- plot_pie_noleg(df_combined, "Dimi", "Distance")
p4 <- plot_pie_noleg(df_combined, "Ramon", "Distance")

# 7. In Patchwork-Layout mit relativen Grössen
# Oben: Emissionen → Breite gemäss em_ratio
# Unten: Distanzen → Breite gemäss km_ratio

((p1 + p2 + plot_layout(widths = c(1, em_ratio))) /
 (p3 + p4 + plot_layout(widths = c(1, km_ratio))))

```

<br><br>

@fig-co2-by-Transportmode illustrates the transport-related CO₂e emissions (in kg) ofDimi and Ramon over a two-month observation period. Each bar is subdivided by transport mode. A third bar represents the Swiss per capita average, also scaled to two months for comparability.

Over this period, Dimi travelled a total of 6,878 km, resulting in 230 kg CO₂e emissions, while Ramon travelled 2,730 km, causing 185 kg CO₂e emissions. Although Dimi covered more than twice the distance, his emissions were only moderately higher – reflecting differences in transport mode choices.

For Dimi, the top three CO₂e sources were car, train, and bus travel. As shown in @tbl-dimi-co2,
he travelled 1,361 km by car, causing 151 kg CO₂e, followed by 4,938 km by train resulting in 54 kg CO₂e, and 197 km by bus leading to 23 kg CO₂e.
These three modes account for over 97% of his emissions. The relatively high value for train travel is largely explained by a long-distance journey to Spain (~3,100 km).

```{r}
#| label: tbl-dimi-co2
#| tbl-cap: CO₂ emissions by transport mode for Dimi over the two-month period.

tibble::tibble(
  `Transport Mode` = c("Car", "Train", "Bus"),
  `Distance (km)` = c(1361, 4938, 197),
  `CO₂ Emissions (kg)` = c(151, 54, 23)
) %>%
  kable(
    col.names = c("Transport Mode", "Distance (km)", "CO₂e Emissions (kg)"),
    align = "lrr"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE
  )

```

<br><br>

For Ramon, the largest contributors were bus, car, and train. As shown in @tbl-ramon-co2,
he travelled 1,188 km by bus, which caused 135 kg CO₂e, followed by 305 km by car (34 kg CO₂e) and 975 km by train (11 kg CO₂e).

```{r}
#| label: tbl-ramon-co2
#| tbl-cap: CO₂ emissions by transport mode for Ramon over the two-month period.

tibble::tibble(
  `Transport Mode` = c("Bus", "Car", "Train"),
  `Distance (km)` = c(1188, 305, 975),
  `CO₂ Emissions (kg)` = c(135, 34, 11)
) %>%
  kable(
    col.names = c("Transport Mode", "Distance (km)", "CO₂e Emissions (kg)"),
    align = "lrr"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE
  )

```

<br><br>

This shows that Ramon’s emissions are strongly shaped by his reliance on buses, while train and car played a secondary role.

The bar labelled Swiss Average corresponds to a reported annual mobility footprint of 2.4 tonnes CO₂e per person in Switzerland [@nathani2022umwelt], scaled down to 400 kg CO₂e for a two-month period.

The dashed black line marks the 1.5°C-compatible emissions threshold, calculated at 164 kg CO₂e per person over two months. This value is derived by first converting the estimated Swiss annual mobility footprint of 2.4 tonnes CO₂e per capita [@nathani2022umwelt] into a two-month equivalent (400 kg CO₂e), and then applying the 59% reduction required in 1.5°C-aligned scenarios, as reported by the IPCC [@IPCC_2022_WGIII_Ch_10]. The resulting value reflects the emissions ceiling for transport per person during a two-month period, if Switzerland were to follow a trajectory consistent with limiting global warming to 1.5°C.

As shown in the graphic, both individuals slightly exceed this threshold – Dimi with 230 kg CO₂e and Ramon with 185 kg CO₂e – but remain significantly below the current national average of 400 kg CO₂e. A more detailed assessment of this benchmark and its implications will be addressed in the discussion section.

```{r}
#| label: fig-co2-by-Transportmode
#| fig-cap: "CO₂e Emissions by Transport Mode"

# 1. Alle Transportmittel erfassen
all_labels <- sort(unique(emissions_combined_filtered$activity_label))

# 2. Dimis CO₂-Werte sortieren (auch 0-Werte einbeziehen)
dimi_sorted_co2 <- emissions_combined_filtered %>%
  filter(person == "Dimi") %>%
  group_by(activity_label) %>%
  summarise(total_co2 = sum(total_emission_kg), .groups = "drop") %>%
  right_join(tibble(activity_label = all_labels), by = "activity_label") %>%
  mutate(total_co2 = replace_na(total_co2, 0)) %>%
  arrange(total_co2) %>%
  pull(activity_label)

# 3. CO₂-Daten aggregieren
plot_data_co2 <- emissions_combined_filtered %>%
  group_by(person, activity_label) %>%
  summarise(total_co2 = sum(total_emission_kg), .groups = "drop") %>%
  mutate(activity_label = factor(activity_label, levels = dimi_sorted_co2))

# 4. Farben (zufällig aber stabil)
set.seed(123)
base_colors <- c(brewer.pal(12, "Set3"),
                 "#999999", "#CCCCCC", "#666666", "#FFB6C1", "#87CEFA")
transport_colors <- setNames(sample(base_colors, length(all_labels)), all_labels)

# 5. Swiss Average hinzufügen
dimi_sorted_co2_extended <- c(dimi_sorted_co2, "Swiss Average")
swiss_row <- tibble(person = "Swiss Average", activity_label = "Swiss Average", total_co2 = 400)

plot_data_extended <- bind_rows(plot_data_co2, swiss_row) %>%
  mutate(activity_label = factor(activity_label, levels = dimi_sorted_co2_extended))

# 6. Farben ergänzen
transport_colors_extended <- transport_colors
transport_colors_extended["Swiss Average"] <- "#d60000"

# 7. Summen & Plot
totals_extended <- plot_data_extended %>%
  group_by(person) %>%
  summarise(total = sum(total_co2)) %>%
  mutate(label = paste0(round(total), " ~kg~CO[2]"))

text_offset_extended <- 0.05 * max(totals_extended$total)

ggplot(plot_data_extended, aes(x = person, y = total_co2, fill = activity_label)) +
  geom_bar(stat = "identity") +
  geom_text(
    data = totals_extended,
    aes(x = person, y = total + text_offset_extended, label = label),
    inherit.aes = FALSE,
    size = 5,
    parse = TRUE
  ) +
  geom_hline(yintercept = 164, linetype = "dashed", color = "black", linewidth = 1) +
  labs(
    x = "Person",
    y = expression("CO"[2] * " Emissions (kg)"),
    fill = "Transport Mode"
  ) +
  scale_fill_manual(values = transport_colors_extended, na.translate = FALSE, drop = FALSE) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title = element_text(size = 14),
    plot.title = element_text(hjust = 0.5, size = 16),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13)
  )

```

<br><br>

# Discussion

<br><br>

## Potential Challenges

A potential challenge in this study is missing or incomplete GPS data from Google Timeline. These gaps may arise due to GPS inaccuracies, device settings, or network issues, leading to incomplete movement records. To address this, missing data will be extrapolated using interpolation techniques, such as linear interpolation or time-weighted averages, ensuring continuity in movement trends. <span style="color:red;">wotsch das würkli no mache?</span> 

Another issue is the misclassification of transportation modes, where Google Timeline may inaccurately categorize movement (e.g., mistaking walking for biking). To improve data reliability, transportation modes will be cross-verified using movement speed, timestamps, and contextual information, allowing for necessary corrections and more accurate mobility analysis.

<br><br>

# Disclaimer

This paper made use of OpenAI's ChatGPT [@openai2024chatgpt] as a supportive tool throughout the research and writing process. ChatGPT was employed to assist in the following areas:
•	Providing language suggestions for clarity and flow
•	Assisting with the structuring and wording of academic content
•	Supporting the formulation and verification of R and Quarto code
All content generated with the help of ChatGPT was critically reviewed, verified, and edited by the author(s) to ensure accuracy, relevance, and academic integrity.

<br><br>


